--- 
title: "Climate downscaling"
author: "Brendan Casey"
#supervisor: 
site: bookdown::bookdown_site
output: 
  md_document:
    variant: gfm
    toc: TRUE
    toc_depth: 2
#bibliography: library.bib
#citation-style: ../../MendeleyBib/CitationStyles/nature.csl
#citation-style: https://www.zotero.org/styles/elsevier-harvard
documentclass: book
link-citations: true
github-repo: https://github.com/bgcasey/climate_downscaling
description: "This is document walks through the analyses I performed for this research."


---


# Overview

```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 100px;
}
```


```{r GlobalOptions, echo=FALSE}
options(bookdown.render.file_scope = FALSE, knitr.duplicate.label = "allow") 
knitr::opts_chunk$set(cache.path = "5_cache/")
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```


```{r setup, include=FALSE, cache=FALSE}
#Set root directory to R project root
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```



```{css, echo=FALSE}
# set max height of code chunks using the following css styling

pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 100px;
}
```


```{r echo=FALSE, out.width = '100%'}
library(kableExtra)
library(knitr)
```




# iButton Data


## Prepare data

Use `1_code/r_notebooks/ibutton_data_prepare.Rmd`.

```{r child = '1_code/r_notebooks/ibutton_data_prepare.Rmd'}
```


## Create spatial objects {-}

Use `1_code/r_notebooks/ibutton_data_xy.Rmd`.

```{r child = '1_code/r_notebooks/ibutton_data_xy.Rmd'}
```


## Quality control {-}

Use `1_code/r_notebooks/3_ibutton_qualityControl.Rmd`.

```{r child = '1_code/r_notebooks/3_ibutton_qualityControl.Rmd'}
```


--- 

# ClimateNA

Use `1_code/r_notebooks/get_climateNA.Rmd`

```{r child = '1_code/r_notebooks/4_ibutton_climateNA.Rmd'}
```


--- 

# Covariates


## Spatial

Spatial covariates were extracted using Google Earth Engine's online code editor at [code.earthengine.google.com](http://code.earthengine.google.com/).

Download the Google Earth Engine script by using `git clone https://earthengine.googlesource.com/users/bgcasey/climate_downscaling` into your working directory. 

Use `1_code/r_notebooks/5_covariates_gee_spatial.Rmd`.

```{r child = '1_code/r_notebooks/5_covariates_gee_spatial.Rmd'}
```

--- 

# Modelling


## Data exploration and visualization

Use `1_code/r_notebooks/modelling_data_exploration.Rmd`.
```{r child = '1_code/r_notebooks/modelling_data_exploration.Rmd'}
```


## Model selection

Use `1_code/r_notebooks/modelling_model_selection.Rmd`.
```{r child = '1_code/r_notebooks/modelling_model_selection.Rmd'}
```


## Offset raster

Use `1_code/r_notebooks/modelling_offset_raster.Rmd`.
```{r child = '1_code/r_notebooks/modelling_offset_raster.Rmd'}
```


## Validate

Use `1_code/r_notebooks/modelling_validate_model.Rmd`.

```{r child = '1_code/r_notebooks/modelling_validate_model.Rmd'}
```




# References


<div id="refs"></div>



