---
title: "Workflow for Refining ClimateNA Temperature Predictions"
subtitle: "BRT model results"
author: "Brendan Casey"
site: bookdown::bookdown_site
output: 
    bookdown::gitbook:
      split_by: "chapter"
bibliography: [../library.bib, ../packages.bib]
link-citations: true
# csl: ecology.csl
citation-style: https://www.zotero.org/styles/taylor-and-francis-apa
---

Here are summaries of the boosted regression trees we produced that predict differences between ClimateNA temperature predictions and micro-climate conditions. Temperature offset layers and project spatial data can be viewed at <https://bgcasey.users.earthengine.app/view/climateoffsets>. Code and a description of the project's workflow can be found at <https://github.com/bgcasey/climate_downscaling>.

```{r setup, include=FALSE, cache=FALSE}
#Set root directory to R project root
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
# 
# setwd(rprojroot::find_rstudio_root_file())
```

```{r include=FALSE}
library(knitr)
library(kableExtra)
library(xlsx)
library(tinytex)
library(rrtable)
library(htmltools)
library(stringr)
library(dismo)
library(huxtable)
library(ggplot2)
library(dplyr)
library(ggstance)
library(jtools)
library(readxl)
library(readr)
```

# Model summaries

## Predictor variables {.unnumbered}

```{r asis=TRUE, message=FALSE, echo=FALSE, warning=FALSE}
# cov_description<-xlsx::read.xlsx("0_data/covariate_list.xlsx", "Sheet2")%>%

cov_description<- read_excel("0_data/covariate_list.xlsx", sheet="Sheet2")  %>%
  # select(1:3)%>%
  filter(Variable=="canopy_height"|
         Variable=="CHILI"|
           Variable=="srad"|
           Variable=="vs"|
           Variable=="northness"|
           Variable=="pr"|
           Variable=="tpi"|
           Variable=="TWI"
           )

kable(cov_description,
  align = c("l","l","r", "r"), caption = 'Spatial covariates included in top models.')%>% 
  # column_spec(column = 3, width = "50em") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 10, full_width = F)%>% 
  scroll_box(width = "95%")
  # kable_styling(font_size = 6, position = "center", full_width = T)
```

## gbm parameters {.unnumbered}

```{r message=FALSE, echo=FALSE, warning=FALSE}
m_param<-read_csv("3_output/tables/brt_params.csv")

# m_param<-m_param%>%mutate(across(where(is.numeric), round, 2))

kable(m_param, caption = 'Tuned BRT parameters.')%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 10, full_width = F)%>% 
  scroll_box(width = "95%")
```

## Variable importance {.unnumbered}

```{r message=FALSE, echo=FALSE}
load("3_output/tables/varimp_brt_seasons.rData")

varimp_brt_seasons<-varimp_brt_seasons%>%mutate(across(where(is.numeric), round, 2))

df<-varimp_brt_seasons
df2 <- data.frame(t(df[-1]))
colnames(df2) <- df[, 1]
df2$model_name<-rownames(df2)
df2<-df2%>%select(c(model_name, srad:season_4))
rownames(df2) <- NULL


kable(df2, caption = 'Variable importance in BRTs.')%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 10, full_width = F)%>% 
  scroll_box(width = "95%")

```

## CV stats {.unnumbered}

```{r message=FALSE, echo=FALSE}
load("3_output/tables/cvstats_brt_seasons.rData")

cvstats_brt_seasons<-cvstats_brt_seasons%>%
  mutate(across(2:5, as.numeric))%>%
  mutate(across(where(is.numeric), round, 2))




# cvstats_brt_seasons$model_name<-str_replace_seasons(cvstats_brt_seasons$model_name, "_tuned", "")

kable(cvstats_brt_seasons, caption = 'BRT cv stats.')%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 10, full_width = F)%>% 
  scroll_box(width = "95%")

```

```{r include=FALSE}
load("2_pipeline/store/models/brt_minTemp_allSeasons.rData")
load("2_pipeline/store/models/brt_maxTemp_allSeasons.rData")
load("2_pipeline/store/models/brt_meanTemp_allSeasons.rData")


load("2_pipeline/store/models/brt_minTemp_summer.rData")
load("2_pipeline/store/models/brt_minTemp_spring.rData")
load("2_pipeline/store/models/brt_minTemp_winter.rData")
load("2_pipeline/store/models/brt_minTemp_fall.rData")

load("2_pipeline/store/models/brt_maxTemp_summer.rData")
load("2_pipeline/store/models/brt_maxTemp_spring.rData")
load("2_pipeline/store/models/brt_maxTemp_winter.rData")
load("2_pipeline/store/models/brt_maxTemp_fall.rData")

load("2_pipeline/store/models/brt_meanTemp_winter.rData")
load("2_pipeline/store/models/brt_meanTemp_spring.rData")
load("2_pipeline/store/models/brt_meanTemp_summer.rData")
load("2_pipeline/store/models/brt_meanTemp_fall.rData")
```

# Partial dependance plots and interactions

## Mean temperature difference {.unnumbered}

### all seasons {.unnumbered}

#### Interactions {.unnumbered}

```{r message=FALSE, echo=FALSE}
int<-gbm.interactions(brt_meanTemp_allSeasons)

kable(int$interactions)%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 10, full_width = F)%>% 
  scroll_box(width = "95%")
```

#### Partial dependance plots {.unnumbered}

```{r  fig.height=12, echo=FALSE, out.width="85%"}
gbm.plot(brt_meanTemp_allSeasons, plot.layout = c(4,2), write.title = FALSE, smooth=TRUE)
```

### Winter {.unnumbered}

#### Interactions {.unnumbered}

```{r message=FALSE, echo=FALSE}
int<-gbm.interactions(brt_meanTemp_winter)

kable(int$interactions)%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 10, full_width = F)%>% 
  scroll_box(width = "95%")
```

#### Partial dependance plots {.unnumbered}

```{r  fig.height=12, echo=FALSE, out.width="85%"}
gbm.plot(brt_meanTemp_winter, plot.layout = c(4,2), write.title = FALSE, smooth=TRUE)
```

### Spring {.unnumbered}

#### Interactions {.unnumbered}

```{r message=FALSE, echo=FALSE}
int<-gbm.interactions(brt_meanTemp_spring)

kable(int$interactions)%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 10, full_width = F)%>% 
  scroll_box(width = "95%")
```

#### Partial dependance plots {.unnumbered}

```{r  fig.height=12, echo=FALSE, out.width="85%"}
gbm.plot(brt_meanTemp_spring, plot.layout = c(4,2), write.title = FALSE, smooth=TRUE)
```

### Summer {.unnumbered}

#### Interactions {.unnumbered}

```{r message=FALSE, echo=FALSE}
int<-gbm.interactions(brt_meanTemp_summer)

kable(int$interactions)%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 10, full_width = F)%>% 
  scroll_box(width = "95%")
```

#### Partial dependance plots {.unnumbered}

```{r  fig.height=12, echo=FALSE, out.width="85%"}
gbm.plot(brt_meanTemp_summer, plot.layout = c(4,2), write.title = FALSE, smooth=TRUE)
```

### Fall {.unnumbered}

#### Interactions {.unnumbered}

```{r message=FALSE, echo=FALSE}
int<-gbm.interactions(brt_meanTemp_fall)

kable(int$interactions)%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 10, full_width = F)%>% 
  scroll_box(width = "95%")
```

#### Partial dependance plots {.unnumbered}

```{r  fig.height=12, echo=FALSE, out.width="85%"}
gbm.plot(brt_meanTemp_fall, plot.layout = c(4,2), write.title = FALSE, smooth=TRUE)
```

## Max temperature difference {.unnumbered}

### all seasons {.unnumbered}

#### Interactions {.unnumbered}

```{r message=FALSE, echo=FALSE}
int<-gbm.interactions(brt_maxTemp_allSeasons)

kable(int$interactions)%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 10, full_width = F)%>% 
  scroll_box(width = "95%")
```

#### Partial dependance plots {.unnumbered}

```{r  fig.height=12, echo=FALSE, out.width="85%"}
gbm.plot(brt_maxTemp_allSeasons, plot.layout = c(4,2), write.title = FALSE, smooth=TRUE)
```

### Winter {.unnumbered}

#### Interactions {.unnumbered}

```{r message=FALSE, echo=FALSE}
int<-gbm.interactions(brt_maxTemp_winter)

kable(int$interactions)%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 10, full_width = F)%>% 
  scroll_box(width = "95%")
```

#### Partial dependance plots {.unnumbered}

```{r  fig.height=12, echo=FALSE, out.width="85%"}
gbm.plot(brt_maxTemp_winter, plot.layout = c(4,2), write.title = FALSE, smooth=TRUE)
```

### Spring {.unnumbered}

#### Interactions {.unnumbered}

```{r message=FALSE, echo=FALSE}
int<-gbm.interactions(brt_maxTemp_spring)

kable(int$interactions)%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 10, full_width = F)%>% 
  scroll_box(width = "95%")
```

#### Partial dependance plots {.unnumbered}

```{r  fig.height=12, echo=FALSE, out.width="85%"}
gbm.plot(brt_maxTemp_spring, plot.layout = c(4,2), write.title = FALSE, smooth=TRUE)
```

### Summer {.unnumbered}

#### Interactions {.unnumbered}

```{r message=FALSE, echo=FALSE}
int<-gbm.interactions(brt_maxTemp_summer)

kable(int$interactions)%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 10, full_width = F)%>% 
  scroll_box(width = "95%")
```

#### Partial dependance plots {.unnumbered}

```{r  fig.height=12, echo=FALSE, out.width="85%"}
gbm.plot(brt_maxTemp_summer, plot.layout = c(4,2), write.title = FALSE, smooth=TRUE)
```

### Fall {.unnumbered}

#### Interactions {.unnumbered}

```{r message=FALSE, echo=FALSE}
int<-gbm.interactions(brt_maxTemp_fall)

kable(int$interactions)%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 10, full_width = F)%>% 
  scroll_box(width = "95%")
```

#### Partial dependance plots {.unnumbered}

```{r  fig.height=12, echo=FALSE, out.width="85%"}
gbm.plot(brt_maxTemp_fall, plot.layout = c(4,2), write.title = FALSE, smooth=TRUE)
```

## Min temperature difference {.unnumbered}

### All seasons {.unnumbered}

#### Interactions {.unnumbered}

```{r message=FALSE, echo=FALSE}
int<-gbm.interactions(brt_minTemp_allSeasons)

kable(int$interactions)%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 10, full_width = F)%>% 
  scroll_box(width = "95%")
```

#### Partial dependance plots {.unnumbered}

```{r  fig.height=12, echo=FALSE, out.width="85%"}
gbm.plot(brt_minTemp_allSeasons, plot.layout = c(4,2), write.title = FALSE, smooth=TRUE)
```

### Winter {.unnumbered}

#### Interactions {.unnumbered}

```{r message=FALSE, echo=FALSE}
int<-gbm.interactions(brt_minTemp_winter)

kable(int$interactions)%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 10, full_width = F)%>% 
  scroll_box(width = "95%")
```

#### Partial dependance plots {.unnumbered}

```{r  fig.height=12, echo=FALSE, out.width="85%"}
gbm.plot(brt_minTemp_winter, plot.layout = c(4,2), write.title = FALSE, smooth=TRUE)
```

### Spring {.unnumbered}

#### Interactions {.unnumbered}

```{r message=FALSE, echo=FALSE}
int<-gbm.interactions(brt_minTemp_spring)

kable(int$interactions)%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 10, full_width = F)%>% 
  scroll_box(width = "95%")
```

#### Partial dependance plots {.unnumbered}

```{r  fig.height=12, echo=FALSE, out.width="85%"}
gbm.plot(brt_minTemp_spring, plot.layout = c(4,2), write.title = FALSE, smooth=TRUE)
```

### Summer {.unnumbered}

#### Interactions {.unnumbered}

```{r message=FALSE, echo=FALSE}
int<-gbm.interactions(brt_minTemp_summer)

kable(int$interactions)%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 10, full_width = F)%>% 
  scroll_box(width = "95%")
```

#### Partial dependance plots {.unnumbered}

```{r  fig.height=12, echo=FALSE, out.width="85%"}
gbm.plot(brt_minTemp_summer, plot.layout = c(4,2), write.title = FALSE, smooth=TRUE)
```

### Fall {.unnumbered}

#### Interactions {.unnumbered}

```{r message=FALSE, echo=FALSE}
int<-gbm.interactions(brt_minTemp_fall)

kable(int$interactions)%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 10, full_width = F)%>% 
  scroll_box(width = "95%")
```

#### Partial dependance plots {.unnumbered}

```{r  fig.height=12, echo=FALSE, out.width="85%"}
gbm.plot(brt_minTemp_fall, plot.layout = c(4,2), write.title = FALSE, smooth=TRUE)
```

# References
<div id="refs"></div>


