---
title: "R Notebook"
output: html_notebook
---

```{r setup, include=FALSE, cache=FALSE}
#Set root directory to R project root
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```


# Setup

##### Load libraries {-}

```{r eval=FALSE}
library(sf)
library(tmap)
```


##### Import ibutton data {-}

```{r eval=FALSE}
hills<-read.csv(file="0_data/external/Hills_iButton_Data_Combined_Corrected_for_Deployment_no_extremes_Apr_27.csv")

RIVR<-read.csv(file="0_data/external/iButtons_RIVR_combined_April7_2022_no_extremes.csv")
```


# Combine ibutton data

##### Create spatial data frame of iButtopn locations {-}

```{r eval=FALSE}
RIVR_xy<-RIVR%>%
  dplyr::select(c(Project, Site_StationKey, Date_deplo, Lat, Long))%>%
  dplyr::distinct()

RIVR_xy<-st_as_sf(RIVR_xy, coords=c("Long","Lat"), crs=4326)

# save as spatial data frame
save(RIVR_xy, file="0_data/manual/spatial/RIVR_xy.rData")

# save as shapefile
st_write(RIVR_xy, "0_data/manual/spatial/RIVR_xy.shp")
````

##### Identify study area {-}

```{r eval=FALSE}
study_area<-st_bbox(RIVR_xy)


```




# Plot
# m<-tm_shape(alberta)+tm_borders()+tm_fill(col = "#fddadd")+
#   #tm_polygons(col=NA, border.col="black")+
#   tm_layout(frame=FALSE)+
#   tm_legend(outside=TRUE, frame=FALSE)+
m<-tm_shape(RIVR_xy)+
    tm_symbols(col = "#2e3c78", border.lwd = 0, size = .3, alpha=.3, title.shape="station location", legend.format = list(text.align="right", text.to.columns = TRUE))
  #tm_legend(position=c("left", "top"), frame=TRUE)
m

tmap_save(m, "3_output/maps/RIVR_xy.png", asp=0, height = 5, width = 6, outer.margins=c(0,0,0,.08))


```








