```{r setup, include=FALSE, cache=FALSE}
#Set root directory to R project root
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

## Setup {.unnumbered}

**Install `ClimateNAr`**

Get the latest ClimateBC/NA r package by registering at a <https://register.climatena.ca/>. Follow the instructions to install the `ClimateNAr` package.

```{r eval=FALSE}
library(ClimateNAr)
library(dplyr)
library(sf)
library(tidyr)
```

## Create input file {.unnumbered}

ClimateNAr requires a properly formatted .csv input file that includes the lat, long, and elevation of each station.

1.  Bring in iButton data\*\*

```{r eval=FALSE}
# load("0_data/manual/iButton_data/ibuttons_complete_monthly.rData")
load("0_data/manual/iButton_data/spatial/ss_xy.rData")

# add month and year columns to the ibutton spatial dataframe
# ibutton<-ss_xy%>%
#       left_join(ibuttons_complete_monthly)%>%
#       select(Project, Site_StationKey, Month, Year)
```

2.  Bring in the elevation data extracted via Google Earth Engine.

```{r eval=FALSE}
elev<-read.csv("0_data/manual/gee_tables/ibutton_terrain.csv")
```

3.  Join elevation data to the iButton spatial data frame.

```{r eval=FALSE}
climateNA_input<- ss_xy%>%
    left_join(elev, by=c('Site_StationKey'='St_SttK'))%>%
    as.data.frame()%>%
    select(Site_StationKey, Lat, Long, elevation)%>%
    drop_na()

write.csv(climateNA_input, file="0_data/manual/iButton_data/climateNA_input.csv")  

```


## Extract CLimateNA monthly summaries

Due to throttling, ClimateNAr's server allows no more than 100 entries (locations). Break the data into mulyiple smaller files if necessary. 

```{r eval=FALSE}
#test with only one location
latLonEl<- c(49.09203,-110.7385, 882)

clm_1 <- ClimateNA_API(ClimateBC_NA='NA',
latLonEl,period=2008,MSY='M');

clm_2 <- ClimateNA_API(ClimateBC_NA='NA',
latLonEl,period=2015,MSY='M');

clm <- ClimateNA_API2 (ClimateBC_NA='NA', inputFile="0_data/manual/iButton_data/climateNA_input.csv",
period='Normal_1961_1990.nrm',MSY='Y');


test<-function(exe ="ClimateNA_v7.30.exe", wkDir="climateNA/", period =
'Normal_1961_1990.nrm', MSY = 'Y', inputFile="0_data/manual/iButton_data/climateNA_input.csv",outputFile= "0_data/manual/iButton_data/climateNA_output.csv")

  
library(ClimateNAr, lib.loc="library")
wkDir = 'library/ClimateNAr'
#exe <- "ClimateNA_v7.30.exe"
inputFile = "0_data/manual/iButton_data/climateNA_input.csv"
outputFile = "0_data/manual/iButton_data/climateNA_output.csv"
period = 'Normal_1961_1990.nrm'
ClimateNA_cmdLine(exe="ClimateNA_v7.30.exe", wkDir='library/ClimateNAr', period='Normal_1961_1990.nrm', MSY='Y',inputFile="0_data/manual/iButton_data/climateNA_input.csv", outputFile="0_data/manual/iButton_data/climateNA_output.csv")



inputFile = 'C:\\Climatena_v730\\InputFiles\\na50k.asc'
outputFile = 'C:\\Climatena_v730\\test\\'
period = 'Normal_1961_1990.nrm'
ClimateNA_cmdLine(exe,wkDir,period,MSY='SY',inputFile, outputFile)  
  
  
  
  
  
  

#Define min and max year
maxYear<-2018
minYear<-2020



for (y in seq(minYear,maxYear)){
  #loop through for each year# it must be the home directory of ClimateNA
  exe <- c
  #inputFile = paste0(data,"temp_input.csv") #find way to change exported file location earlier into format needed here?
  #outputFile = paste0(data,"temp_input_Year_",y,"MP.csv")
  # input = "0_data/manual/iButton_data/climateNA_input.csv" 
  #outputFile = paste0("0_data/climateNA/temp_input_Year_",y,"MP.csv")
  yearPeriod = paste0('\"Year_',y,'.ann\"')
  paste0("clm_", y)<-ClimateNA_API2(ClimateBC_NA='NA',  inputFile="0_data/manual/iButton_data/climateNA_input.csv" , period=yearPeriod, MSY='Y') 
  #dat <- read.csv('C:/Users/kimorris/ClimateNA_v730/ibutton/temp_output.csv'); head(dat) 
}
  



#Inputs
#ClimateNA executable
c<-"ClimateNA_v7.30.exe"
#input directory
#e.g."/C:\\Users\\kimorris\\ClimateNA_v730\\ibutton\\"
#data<-'/C:\\Users\\kimorris\\ClimateNA_v730\\ibutton\'


maxYear<-2018
minYear<-2020


for (y in seq(minYear,maxYear)){
  #loop through for each year
  setwd(climNA_dir);getwd() # it must be the home directory of ClimateNA
  exe <- c
  #inputFile = paste0(data,"temp_input.csv") #find way to change exported file location earlier into format needed here?
  #outputFile = paste0(data,"temp_input_Year_",y,"MP.csv")
  inputFile = paste0("/C:\\Users\\kimorris\\ClimateNA_v730\\ibutton\temp_input.csv") #find way to change exported file location earlier into format needed here?
  outputFile = paste0("/C:\\Users\\kimorris\\ClimateNA_v730\\ibutton\temp_input.csv","temp_input_Year_",y,"MP.csv")
  yearPeriod = paste0('/Year_',y,'.ann')
  system2(exe,args= c('/Y', yearPeriod, inputFile, outputFile)) 
  #dat <- read.csv('C:/Users/kimorris/ClimateNA_v730/ibutton/temp_output.csv'); head(dat) 
}









y=2020
yearPeriod = paste0('\"Year_',y,'.ann\"')

paste0("clm_", y)<-ClimateNA_API2(ClimateBC_NA='NA', MSY='Y', inputFile=input, period=yearPeriod, inputFile=input)


ib_clim<-NULL
for (y in seq(minYear,maxYear)){
  file<-paste0("/0_data/climateNA/","temp_input_Year_",y,"MP.csv")
  print(file)
  t<-read.csv(file)
  print(head(t))
  tt<-subset(t,select=c("ID1","ID2","Latitude","Longitude","Elevation", "Tmax01","Tmax02","Tmax03","Tmax04","Tmax05","Tmax06","Tmax07","Tmax08","Tmax09","Tmax10","Tmax11","Tmax12"))
  print(head(tt))
  t_long <- melt(setDT(tt), id.vars = c("ID1","ID2","Latitude","Longitude","Elevation"), variable.name = "month")
  print(head(t_long))
  #select only needed columns
  t_long2<-subset(t_long,select=c("ID1","month","value"))
  #parse month from Tmax01 etc
  t_long2$month<-str_sub(t_long2$month, start= -2)
  #name value column ClimNATmax
  colnames(t_long2)<-c("group","month","climNA_tmax")
  #remove leading 0's on month
  t_long2$month<-sub("^0+", "",t_long2$month)
  #head(t_long2)
  #table(t_long2$month)
  #merge with tmax2
  t_tmax<-subset(tmax2,year==y)
  #table(t_tmax$month)
  t_tmax2<-merge(t_tmax,t_long2,by=c("group","month"))
  #head(t_tmax2)
  #table(t_tmax2$month)
  #nrow(t_tmax2)
  #nrow(t_tmax)
  ##merge years as it loops through them
  ib_clim<-rbind(ib_clim,t_tmax2)
}




```
