```{r setup, include=FALSE, cache=FALSE}
#Set root directory to R project root
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

## Setup {.unnumbered}

**Install `ClimateNAr`**

Get the latest ClimateBC/NA r package by registering at a <https://register.climatena.ca/>. Follow the instructions to install the `ClimateNAr` package.

```{r eval=FALSE}
library(ClimateNAr)
library(dplyr)
library(sf)
library(tidyr)
```

Download the ClimateNA desktop application and r package here: https://register.climatena.ca/

Instructions on how use the ClimateNA application can be found here: https://pressbooks.bccampus.ca/fode010notebook/chapter/topic-3-2-the-use-of-climatena-ap-to-generate-point-and-spatial-climate-data/


The climateNA desktop application was designed for PC, but can be used on Mac by using Wine. To get climateNA working on my machine I used a Wineskin (https://github.com/Gcenx/WineskinServer) to create an application wrapper. The wrapper used the WS10Wine64Bit5.20 wine engine. I also installed `vb6run` with winetricks to get the application working. The the wine wrapper should contain the entire `climateNA_v730` folder directory, not just the `ClimateNA_v7.30.exe` file. 

While climateNA can be run via commandline using R, I was not able to get this working on my Mac. I included code for integrating climateNA into the R workflow on PC's ([Extract CLimateNA monthly summaries with R ]). However, the R code is untested. 


## Create input file {.unnumbered}

ClimateNAr requires a properly formatted .csv input file that includes the following headers: ID1, ID2, lat, long, el


1.  Bring in iButton data\*\*

```{r eval=FALSE}
# load("0_data/manual/iButton_data/ibuttons_complete_monthly.rData")
load("0_data/manual/iButton_data/spatial/ss_xy.rData")

# add month and year columns to the ibutton spatial dataframe
# ibutton<-ss_xy%>%
#       left_join(ibuttons_complete_monthly)%>%
#       select(Project, Site_StationKey, Month, Year)
```

2.  Bring in the elevation data extracted via Google Earth Engine.

```{r eval=FALSE}
elev<-read.csv("0_data/manual/gee_tables/ibutton_terrain.csv")
```

3.  Join elevation data to the iButton spatial data frame.

```{r eval=FALSE}
climateNA_input<- ss_xy%>%
    left_join(elev, by=c('Project'='Project', 'Site_StationKey'='St_SttK'))%>%
    as.data.frame()%>%
    select(Project, Site_StationKey, Lat, Long, elevation)%>%
    rename(c(ID1=Project, ID2=Site_StationKey, lat=Lat, long=Long, el=elevation))%>%
    drop_na()

write.csv(climateNA_input, file="0_data/manual/climateNA/input/climateNA_input.csv", row.names = FALSE)  
```

## Extract ClimateNA summaries {-}

I used the ClimateNA GUI to extract monthly climate summaries. There is a weird bug on my mac ClimateNA application where I need to open and save the input .csv in excel before ClimateNA can read it.   

In the desktop application I selected a `Historical Time Series` of `Monthly primary variables (60)`

Climate data is exported back into the rProject as a .csv file in `0_data/manual/climateNA/output/`

### Read in the ClimateNA summaries {-}

```{r eval =FALSE}
cna<-read.csv(file="0_data/manual/climateNA/output/climateNA_input_2015-2021MP.csv")
  
cNA_month_summaries<-cna%>%
  rename(c(Project=ID1, Site_StationKey=ID2))%>%
  dplyr::select(c(Year:Tave12))%>%
  gather(key="Month_a", value="cNA_Tmax", "Tmax01":"Tmax12")%>%
  mutate(Month_a=(gsub("[^0-9.-]", "", Month_a))) %>%
  gather(key="Month_b", value="cNA_Tmin", "Tmin01":"Tmin12")%>%
  mutate(Month_b=(gsub("[^0-9.-]", "", Month_b))) %>%
  gather(key="Month_c", value="cNA_Tave", "Tave01":"Tave12")%>%
  mutate(Month_c=(gsub("[^0-9.-]", "", Month_c))) %>%
  filter(Month_a==Month_b & Month_b==Month_c)%>%
  dplyr::select(Project, Site_StationKey, Year, Month_a, cNA_Tmax, cNA_Tmin, cNA_Tave)%>%
  rename(Month=Month_a)%>%
  arrange(Site_StationKey, Year, Month)

save(cNA_month_summaries, file="0_data/manual/climateNA/processed/cNA_month_summaries.rData")
write.csv(cNA_month_summaries, file="0_data/manual/climateNA/processed/cNA_month_summaries.csv")
```

```{r}
load("0_data/manual/climateNA/processed/cNA_month_summaries.rData")
knitr::kable(head(cNA_month_summaries), "pipe") 
```



## Extract CLimateNA monthly summaries with R {-}

I have not been able to get this working on my Mac, but it should work on PC's (untested).


```{r eval=FALSE}
library(ClimateNAr)
wkDir = '/Applications'
exe <- "ClimateNA_v7_30.app"
inputFile = '0_data/manual/climateNA/input/climateNA_input.csv'
outputFile = '0_data/manual/climateNA/output/climateNA_output.csv'
period = 'Normal_1961_1990.nrm'
ClimateNA_cmdLine(exe, wkDir, period, MSY='Y',inputFile, outputFile)




inputFile = 'C:\\Climatena_v730\\InputFiles\\na50k.asc'
outputFile = 'C:\\Climatena_v730\\test\\'
period = 'Normal_1961_1990.nrm'
ClimateNA_cmdLine(exe,wkDir,period,MSY='SY',inputFile, outputFile)
```

