```{r setup, include=FALSE, cache=FALSE}
#Set root directory to R project root
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

setwd(rprojroot::find_rstudio_root_file())
```

```{r eval=FALSE}
library(tidyr)
library(plyr)
library(dplyr)
library(kableExtra)
library(lubridate)
library(imputeTS)
library(readr)
```

### Import and clean {.unnumbered}

#### RIVR {.unnumbered}

**Load data**

```{r}
RIVR<-read.csv(file="0_data/external/iButton/RIVR/iButtons_RIVR_combined_April7_2022_no_extremes.csv")
```

**Examine data frame**

Count the number of unique iButtons

```{r echo=FALSE, message = FALSE, results="asis"}
# count unique deployments
nrow(distinct(as.data.frame((RIVR$Site_StationKey))))
```

View data

```{r echo=FALSE, message = FALSE, results="asis"}
#RIVR<-read.csv(file="0_data/external/iButton/RIVR/iButtons_RIVR_combined_April7_2022_no_extremes.csv")
knitr::kable(head(RIVR), "pipe") 
```

##### Edit data columns {.unnumbered}

###### Format date-time string {.unnumbered}

Convert date-time string into a `POSIXct` class. This class associates the date time string with an associated time zone. Codes for time zones can be found [here](https://en.wikipedia.org/wiki/List_of_tz_database_time_zones). Else the `as.POSIXct` function defaults to your computer's timezone.

```{r eval=FALSE}
RIVR$Date_Time<-as.POSIXct(RIVR$Date_Time, tz="America/Edmonton")
```

##### Remove unusable data {.unnumbered}

###### Remove months with incomplete data {.unnumbered}

Removing first month of data if it has only 20 days or less.

```{r eval=FALSE}
RIVR_1<- RIVR%>% split(f=.$New_Site_Key) %>% lapply(FUN=function(x){
  if (x %>% filter(month(Date_Time)==month(min(Date_Time))& year(Date_Time)==min(year(Date_Time))) %>%
      summarize(max(Day)-min(Day)) %>% c () < 20 ){
    y<-x %>% filter(month(Date_Time)>month(min(Date_Time))| year(Date_Time)> min(year(Date_Time)))    
    
    return(y)
  }else{
    return(x)
  }
}) 
```

Removing last month of data if it has only 20 days or less.

```{r eval=FALSE}
RIVR_2<-RIVR_1 %>% lapply(FUN=function(x){
  if (x %>% filter(month(Date_Time)==month(max(Date_Time))& year(Date_Time)==max(year(Date_Time))) %>% 
      summarize(max(Day)-min(Day)) %>% c () < 20 ){
    y<-x %>% filter(month(Date_Time)<month(max(Date_Time))| year(Date_Time)< max(year(Date_Time)))    
    
    #y<-x[!(month(x$Date_Time)==month(min(x$Date_Time) & year(x$Date_Time)==min(year(x$Date_Time)))),]
    return(y)
  }else{
    return(x)
  }
  
}) %>% do.call(rbind,.) #working

```

##### Remove pre-deployement data {.unnumbered}

```{r eval=FALSE}
RIVR_3<- RIVR_2%>%
  filter(Date_Time>Date_Time_dpl)%>%
  filter(Date_Time<Date_Time_rtv)

RIVR_cleaned<-RIVR_3
save(RIVR_cleaned, file="2_pipeline/tmp/RIVR_cleaned.rData")
```


##### Identify and remove data from grounded iButtons {.unnumbered}

##### Get daily temperature summaries {.unnumbered}

Create new calculated columns with the mean, max, and min daily temperatures and diurnal range. 

```{r eval=FALSE}
RIVR_3<-RIVR_cleaned %>% 
  mutate(Date=date(Date_Time))%>%
  dplyr::group_by(Site_StationKey,Day,Month,Year,Date, iBt_type)%>%  dplyr::mutate(Temperature=Value) %>%
  dplyr::summarize(Tmax_Day=max(Temperature),Tmin_Day=min(Temperature),Tavg_Day=mean(Temperature), T_diurnal=max(Temperature)-min(Temperature))%>% 
  dplyr::group_by(Site_StationKey,iBt_type,Month,Year) %>% dplyr::filter(!iBt_type=="EXTRA-TOP")%>%
  arrange(Site_StationKey, Date)

RIVR_dailys<-RIVR_3

save(RIVR_dailys, file="2_pipeline/store/RIVR_dailys.rData")
```



```{r echo=FALSE, message = FALSE, results="asis"}
load("2_pipeline/store/RIVR_dailys.rData")

knitr::kable(head(RIVR_dailys), "pipe") 
```





#### HILL {.unnumbered}

**Load data**

```{r}
hills<-read.csv(file="0_data/external/iButton/Hills/Hills_iButton_Data_Combined_Corrected_for_Deployment_no_extremes_Apr_27.csv")
```

**Examine data frame**

Count the number of unique iButtons

```{r echo=FALSE, message = FALSE, results="asis"}
# count unique deployments
nrow(distinct(as.data.frame((hills$Site_StationKey))))
```

View data

```{r echo=FALSE, message = FALSE, results="asis"}
#RIVR<-read.csv(file="0_data/external/iButton/RIVR/iButtons_RIVR_combined_April7_2022_no_extremes.csv")
knitr::kable(head(hills), "pipe") 
```

##### Edit data columns {.unnumbered}

###### Format date-time string {.unnumbered}

Convert date-time string into a `POSIXct` class. This class associates the date time string with an associated time zone. Codes for time zones can be found [here](https://en.wikipedia.org/wiki/List_of_tz_database_time_zones). Else the `as.POSIXct` function defaults to your computer's timezone.

```{r eval=FALSE}
hills$Date.Time<-as.POSIXct(hills$Date.Time, tz="America/Edmonton")
```

##### Remove unusable data {.unnumbered}

###### Remove months with incomplete data {.unnumbered}

Remove first and last months if they have less than 21 days of data.

```{r eval=FALSE}
hills_1<-hills %>%
  mutate(Value=Temperature,Month=month(Date.Time),Day=day(Date.Time),Year=year(Date.Time),
                Date_Time=Date.Time) %>%  
  ###Split per site_stationkey and remove first and last month if they have less than 20 days
  split(f=.$Site_StationKey) %>% lapply(FUN=function(x){
    if (x %>% filter(month(Date_Time)==month(min(Date_Time))& year(Date_Time)==min(year(Date_Time))) %>%
        summarize(max(Day)-min(Day)) %>% c () < 20 ){
      y<-x %>% filter(month(Date_Time)>month(min(Date_Time))| year(Date_Time)> min(year(Date_Time)))    
      
      return(y)
    }else{
      return(x)
    }
    
  }) %>% lapply(FUN=function(x){
    if (x %>% filter(month(Date_Time)==month(max(Date_Time))& year(Date_Time)==max(year(Date_Time))) %>% 
        summarize(max(Day)-min(Day)) %>% c () < 20 ){
      y<-x %>% filter(month(Date_Time)<month(max(Date_Time))| year(Date_Time)< max(year(Date_Time)))    
      
      #y<-x[!(month(x$Date_Time)==month(min(x$Date_Time) & year(x$Date_Time)==min(year(x$Date_Time)))),]
      return(y)
    }else{
      return(x)
    }
    
  }) %>% do.call(rbind,.) 


hills_cleaned<-hills_1
save(hills_cleaned, file="2_pipeline/tmp/hills_cleaned.rData")

```

##### Remove predeployement data {.unnumbered}

##### Get daily temperature summaries {.unnumbered}

Create new calculated columns with the mean, max, and min daily temperatures.

```{r eval=FALSE}
hills_2<-hills_1 %>%
  mutate(Date=date(Date_Time))%>%
  dplyr::group_by(Site_StationKey,Date) %>% #group by days, month and year
  dplyr::summarize(Tmax_Day=max(Temperature),Tmin_Day=min(Temperature),Tavg_Day=mean(Temperature))%>%
  mutate(Year=year(Date))%>%
  mutate(Month=month(Date))%>%
  mutate(Day=day(Date))%>%
  dplyr::group_by(Site_StationKey,Date) 

hills_dailys<-hills_2

save(hills_dailys, file="2_pipeline/store/hills_dailys.rData")
```

```{r echo=FALSE, message = FALSE, results="asis"}
load("2_pipeline/store/hills_dailys.rData")

knitr::kable(head(hills_dailys), "pipe") 
```


#### Wood {-} 

Data from Wood, Wendy H; Marshall, Shawn J; Fargey, Shannon E; Whitehead, Terri L (2017): Daily temperature data from the Foothills Climate Array Mesonet, Canadian Rocky Mountains, 2005-2010. PANGAEA, https://doi.org/10.1594/PANGAEA.880611. 

```{r eval=FALSE}
wood<-read.delim(file="0_data/external/iButton/WoodEtAl/Wood_etal_2017.tab",  header=T, sep="\t", skip = 23)
```


#### Alex {-}

```{r eval=FALSE}

# read csvs
alex_2019<-read_csv(file="0_data/external/iButton/alex/combined/Combined2019TempData.csv")
alex_2020_2021<-read_csv(file="0_data/external/iButton/alex/combined/Combined2020_2021TempData.csv")


# need to add a time column to alex_2020_2021
## read and rbind csvs



# read in the individidual 2020-2021 csvs
filenames=list.files(path="0_data/external/iButton/alex/individual/2020_2021", full.names=TRUE)
myMergedData <- 
    read_csv(filenames) %>% 
    bind_rows()
colnames(myMergedData)[1] <- "date_time"

# fix date and time fields 
alex_2020_2021a<-myMergedData%>%
  mutate(date_time=gsub("\\.","",myMergedData$date_time))%>%
  mutate(site_id=sub("\\(.*","",myMergedData$site_id))%>% # remove entore string after (
  mutate(site_id=gsub(" ", "", site_id, fixed = TRUE))%>% # remove white space
  mutate(site_id=gsub("(?<=\\d)(?=\\D)|(?<=\\D)(?=\\d)", "_", site_id, perl = TRUE)) %>% # add _ between digits and letters
  drop_na(c(date_time, date_time_deployed))%>%
  mutate(date_time_2=dmy_hm(date_time))%>%
  # mutate(date_time_3=replace_na(list(date_time_2), 0))%>%
  mutate(date_time_3 = ifelse(is.na(date_time_2), dmy_hms(date_time), date_time_2))%>%
  mutate(date_time_corrected= as_datetime(date_time_3))%>%
  # mutate(date_time_all=coalesce(date_time_2, date_time_3))%>%
  select(-c(date_time_2, date_time_3))%>%
  mutate(month=month(date_time_corrected))%>%
  mutate(day=day(date_time_corrected))%>%
  mutate(year=year(date_time_corrected))%>%
  mutate(date_time_deployed_1=dmy_hm(date_time_deployed))%>%
  mutate(date_time_deployed_2 = ifelse(is.na(date_time_deployed_1), dmy_hms(date_time_deployed), date_time_deployed_1))%>%
  mutate(date_time_deployed_corrected= as_datetime(date_time_deployed_2))%>%
  select(-c(date_time_deployed_2, date_time_deployed_1))%>%
  # mutate(date_time_deployed_1=dmy_hms(date_time_deployed))%>%
  arrange(temp_id, date_time_deployed_corrected)%>%
  filter(date_time_deployed_corrected<date_time_corrected)%>%
  distinct()
 
unique(alex_2020_2021$site_id)
# there are a lot more site id's here than on Alexs plot locations spreadsheet. 557 vs 97. Missing somthing? Some of the ids in the locations and elevations spreadsheet are missing from the data (e.g. 1_WA, 3_C)


alex_2020_2021_fin<-read_csv(file="0_data/external/iButton/alex/combined/Combined2020_2021TempData.csv")


test<-alex_2020_2021[c(4:5)]%>%distinct()
test2<-alex_2020_2021_fin%>%
  select(site_id, plot_id)%>%distinct()
test3<-alex_2019%>%
      select(8)%>%
      distinct()
```


### Combine datasets {.unnumbered}

#### **Bind dataframes**

```{r eval=FALSE}

ibuttons<-bind_rows("RIVR"=RIVR_dailys,"HILLS"=hills_dailys,.id="Project")

save(ibuttons, file="2_pipeline/tmp/ibuttons.rData")
```

