```{r setup, include=FALSE, cache=FALSE}
#Set root directory to R project root
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

setwd(rprojroot::find_rstudio_root_file())
```

```{r eval=FALSE}
library(tidyr)
library(plyr)
library(dplyr)
library(kableExtra)
library(lubridate)
library(imputeTS)
library(readr)
```

### Import and clean {.unnumbered}

#### RIVR {.unnumbered}

**Load data**

```{r}
RIVR<-read.csv(file="0_data/external/iButton/RIVR/iButtons_RIVR_combined_April7_2022_no_extremes.csv")
```

**Examine data frame**

Count the number of unique iButtons

```{r echo=FALSE, message = FALSE, results="asis"}
# count unique deployments
nrow(distinct(as.data.frame((RIVR$Site_StationKey))))
```

View data

```{r echo=FALSE, message = FALSE, results="asis"}
#RIVR<-read.csv(file="0_data/external/iButton/RIVR/iButtons_RIVR_combined_April7_2022_no_extremes.csv")
knitr::kable(head(RIVR), "pipe") 
```

##### Edit data columns {.unnumbered}

###### Format date-time string {.unnumbered}

Convert date-time string into a `POSIXct` class. This class associates the date time string with an associated time zone. Codes for time zones can be found [here](https://en.wikipedia.org/wiki/List_of_tz_database_time_zones). Else the `as.POSIXct` function defaults to your computer's timezone.

```{r eval=FALSE}
RIVR$Date_Time<-as.POSIXct(RIVR$Date_Time, tz="America/Edmonton")
```

##### Remove pre-deployement data {.unnumbered}

```{r eval=FALSE}
RIVR_2<- RIVR%>%
  filter(Date_Time>Date_Time_dpl)%>%
  filter(Date_Time<Date_Time_rtv)
```


##### Get daily temperature summaries {.unnumbered}

Create new calculated columns with the mean, max, and min daily temperatures and diurnal range. 

```{r eval=FALSE}
RIVR_dailys<-RIVR_2 %>% 
  mutate(Date=date(Date_Time))%>%
  dplyr::group_by(Site_StationKey,Day,Month,Year,Date, iBt_type)%>%  dplyr::mutate(Temperature=Value) %>%
  dplyr::summarize(Tmax_Day=max(Temperature),Tmin_Day=min(Temperature),Tavg_Day=mean(Temperature))%>%
  dplyr::group_by(Site_StationKey,iBt_type,Month,Year) %>% dplyr::filter(!iBt_type=="EXTRA-TOP")%>%
  arrange(Site_StationKey, Date)

save(RIVR_dailys, file="2_pipeline/store/RIVR_dailys.rData")
```



```{r echo=FALSE, message = FALSE, results="asis"}
load("2_pipeline/store/RIVR_dailys.rData")

knitr::kable(head(RIVR_dailys), "pipe") 
```


#### HILL {.unnumbered}

**Load data**

```{r}
hills<-read.csv(file="0_data/external/iButton/Hills/Hills_iButton_Data_Combined_Corrected_for_Deployment_no_extremes_Apr_27.csv")
```

**Examine data frame**

Count the number of unique iButtons

```{r echo=FALSE, message = FALSE, results="asis"}
# count unique deployments
nrow(distinct(as.data.frame((hills$Site_StationKey))))
```

View data

```{r echo=FALSE, message = FALSE, results="asis"}
#RIVR<-read.csv(file="0_data/external/iButton/RIVR/iButtons_RIVR_combined_April7_2022_no_extremes.csv")
knitr::kable(head(hills), "pipe") 
```

##### Edit data columns {.unnumbered}

###### Format date-time string {.unnumbered}

Convert date-time string into a `POSIXct` class. This class associates the date time string with an associated time zone. Codes for time zones can be found [here](https://en.wikipedia.org/wiki/List_of_tz_database_time_zones). Else the `as.POSIXct` function defaults to your computer's timezone.

```{r eval=FALSE}
hills$Date.Time<-as.POSIXct(hills$Date.Time, tz="America/Edmonton")
```


```{r eval=FALSE}
hills_1<-hills %>%
  mutate(Value=Temperature,Month=month(Date.Time),Day=day(Date.Time),Year=year(Date.Time),
                Date_Time=Date.Time) 

```


##### Get daily temperature summaries {.unnumbered}

Create new calculated columns with the mean, max, and min daily temperatures.

```{r eval=FALSE}
hills_dailys<-hills %>%
   mutate(Value=Temperature,Month=month(Date.Time),Day=day(Date.Time),Year=year(Date.Time),
                Date_Time=Date.Time) %>%
  mutate(Date=date(Date_Time))%>%
  dplyr::group_by(Site_StationKey,Date) %>% #group by days, month and year
  dplyr::summarize(Tmax_Day=max(Temperature),Tmin_Day=min(Temperature),Tavg_Day=mean(Temperature))%>%
  mutate(Year=year(Date))%>%
  mutate(Month=month(Date))%>%
  mutate(Day=day(Date))%>%
  dplyr::group_by(Site_StationKey,Date) 

save(hills_dailys, file="2_pipeline/store/hills_dailys.rData")
```

```{r echo=FALSE, message = FALSE, results="asis"}
load("2_pipeline/store/hills_dailys.rData")

knitr::kable(head(hills_dailys), "pipe") 
```


#### Wood {-} 

Data from Wood, Wendy H; Marshall, Shawn J; Fargey, Shannon E; Whitehead, Terri L (2017): Daily temperature data from the Foothills Climate Array Mesonet, Canadian Rocky Mountains, 2005-2010. PANGAEA, https://doi.org/10.1594/PANGAEA.880611. 

```{r eval=FALSE}
wood_dailys<-read.delim(file="0_data/external/iButton/WoodEtAl/Wood_etal_2017.tab",  header=T, sep="\t", skip = 23)%>%
  rename(c(Site_StationKey=Station..FCAID., Date_Time=Date.Time, Tmin_Day=TnTnTn.day.min...C., Tmax_Day=TxTxTx.day.max...C., Tavg_Day=TTT.daily.m...C.))%>%
  select(Site_StationKey, Date_Time, Tmin_Day, Tmax_Day)%>%
  mutate(Date=date(Date_Time))%>%
  mutate(Year=year(Date))%>%
  mutate(Month=month(Date))%>%
  mutate(Day=day(Date))%>%
  select(-Date_Time)

save(wood_dailys, file="2_pipeline/store/wood_dailys.rData")
```


#### Alex {-}

##### Clean data {-}

```{r eval=FALSE}
# read in the individidual 2020-2021 csvs
filenames=list.files(path="0_data/external/iButton/alex/individual/2020_2021", full.names=TRUE)

library(data.table)
myMergedData <- rbindlist(sapply(filenames, fread,simplify = FALSE), idcol = 'filename')
colnames(myMergedData)[2] <- "date_time"

# check the number of site_ids per file
n_site_id<-myMergedData %>%
  group_by(filename) %>%
  mutate(n_id= n_distinct(site_id))%>%
  select(filename, n_id)%>%
  distinct()%>%
  filter(n_id== 1)

# there is a problem with a coupe of the files:
# filename	                                # of site_id's
# 4D0000006E19FA41_101321_Sept20_12PM.csv	  2
# F60000006E206641_101421_Sept23_10AM.csv	  497
# FC0000006E203941_101321_Sept23_10AM.csv	  164

# I'll exclude these for now. 
myMergedData_2<-myMergedData %>%
    group_by(filename) %>%
    mutate(n_id= n_distinct(site_id))%>%
    ungroup()%>%
    filter(n_id == 1)%>%
    select(-n_id)%>%
    select(-filename)

# fix date and time fields 
alex_2020_2021a<-myMergedData_2%>%
  mutate(date_time=gsub("\\.","",myMergedData_2$date_time))%>%
  mutate(site_id=sub("\\(.*","",myMergedData_2$site_id))%>% # remove entore string after (
  mutate(site_id=gsub(" ", "", site_id, fixed = TRUE))%>% # remove white space
  mutate(site_id=gsub("(?<=\\d)(?=\\D)|(?<=\\D)(?=\\d)", "_", site_id, perl = TRUE)) %>% # add _ between digits and letters
  drop_na(c(date_time, date_time_deployed))%>%
  mutate(date_time_2=dmy_hm(date_time))%>%
  # mutate(date_time_3=replace_na(list(date_time_2), 0))%>%
  mutate(date_time_3 = ifelse(is.na(date_time_2), dmy_hms(date_time), date_time_2))%>%
  mutate(date_time_corrected= as_datetime(date_time_3))%>%
  # mutate(date_time_all=coalesce(date_time_2, date_time_3))%>%
  select(-c(date_time_2, date_time_3))%>%
  mutate(month=month(date_time_corrected))%>%
  mutate(day=day(date_time_corrected))%>%
  mutate(year=year(date_time_corrected))%>%
  mutate(date_time_deployed_1=dmy_hm(date_time_deployed))%>%
  mutate(date_time_deployed_2 = ifelse(is.na(date_time_deployed_1), dmy_hms(date_time_deployed), date_time_deployed_1))%>%
  mutate(date_time_deployed_corrected= as_datetime(date_time_deployed_2))%>%
  select(-c(date_time_deployed_2, date_time_deployed_1))%>%
  # mutate(date_time_deployed_1=dmy_hms(date_time_deployed))%>%
  arrange(temp_id, date_time_deployed_corrected)%>%
  filter(date_time_deployed_corrected<date_time_corrected)%>%
  distinct()%>%
  select(-c(date_time_deployed, date_time, temp_id, serial_number, location))%>%
  rename(c(Temperature=Value, Year=year, Day=day, Month=month, Date_Time=date_time_corrected,  Date_Time_dpl=date_time_deployed_corrected, Site_StationKey=site_id ))%>%
  mutate(Date=date(Date_Time))%>%
  mutate(Project="alex")
 

alex_cleaned<-alex_2020_2021a
save(alex_cleaned, file="2_pipeline/tmp/alex_cleaned.rData")


# bring in xy data
alex_xy<-read_csv(file="0_data/external/iButton/alex/2019_and_2020_2021_plot_locations_and_elevations.csv")

xy<-alex_xy%>%
  mutate(site_id=paste(site, plot, sep="_"))

alex_2020_2021_xy<-alex_2020_2021a%>%
  inner_join(xy)


```


##### Get daily temperature summaries {.unnumbered}

Create new calculated columns with the mean, max, and min daily temperatures. 

```{r eval=FALSE}
alex_dailys<-alex_cleaned %>% 
  dplyr::group_by(Site_StationKey,Day,Month,Year,Date)%>%
  dplyr::summarize(Tmax_Day=max(Temperature),Tmin_Day=min(Temperature),Tavg_Day=mean(Temperature))%>% 
  dplyr::group_by(Site_StationKey,Month,Year) %>% 
  arrange(Site_StationKey, Date)

save(alex_dailys, file="2_pipeline/store/alex_dailys.rData")
```


### Combine datasets {.unnumbered}

#### **Bind dataframes**

```{r eval=FALSE}

ibuttons<-bind_rows("RIVR"=RIVR_dailys,"HILLS"=hills_dailys, "alex"=alex_dailys, "wood"=wood_dailys, .id="Project")

save(ibuttons, file="2_pipeline/tmp/ibuttons.rData")
```

